\section*{Acknowledgments}

This thesis was completed as part of the Master of Science programme in Civil Engineering at ETH Zurich undertaken during a year exchange from Imperial College London (Civil Engineering MEng). I would like to express my sincere gratitude to those who contributed to its development.

I am especially grateful to my supervisors at the Chair of Concrete Structures and Bridge Design: Professor Walter Kaufmann, Anna Hodel, and Tena Galkovski. Their technical guidance throughout the project has been invaluable. The opportunity to build on the pre-existing optimisation framework developed within the research group has been a formative experience, and I am appreciative of the time invested in discussing both the engineering principles and the implementation details at our regular meetings.

I am also grateful to the decking manufacturers whose technical data made this comparison possible. In particular, I would like to thank Peter Baker and Barry Rust of Tata Steel, and Joanna Parzyszek of ArcelorMittal, for generously providing product information and Environmental Product Declaration data and for answering my queries on their decking ranges. My thanks also go to Kingspan for making the technical and environmental data for their Multideck range publicly available. The willingness of these companies to share detailed product and EPD information was essential to building the deck database on which this study relies.

Finally, I would like to thank my personal tutor, Professor Adrian Butler, and my exchange coordinator, Dr. Christian Onof, for their encouragement and support throughout the duration of the thesis. The project has spanned a period of sustained focus alongside concurrent coursework commitments, and their patience and advice has been greatly appreciated.

\vspace{12pt}