\section{Discussion}
\label{fifthchapter}

This chapter interprets the results presented in Chapter~\ref{fourthchapter} and situates them within the broader context of sustainable structural design. It first addresses the central and initially counter-intuitive finding---that the composite slab carries the highest embodied carbon of the systems compared, yet composite construction remains widely specified in practice---and then examines the behaviour within and between slab systems, compares the results against the published literature, considers the validity and limitations of the design methodology, and identifies directions for further work.

%==============================================================================
\subsection{Why composite slabs are specified despite higher GWP}
\label{sec:disc_why_composite}
%==============================================================================
The results show unambiguously that, on embodied carbon alone, the composite slab is the least favourable of the five systems across every span, occupancy, and span configuration (Section~\ref{sec:res_office_gwp}). This raises the obvious question of why composite steel--concrete floors are nonetheless among the most common floor systems in multi-storey steel-framed construction. The answer lies in the factors that a per-square-metre embodied-carbon comparison does not capture, and which a designer weighs alongside it.

\paragraph{Construction speed and programme.} Profiled steel decking acts as permanent formwork and a working platform, eliminating the striking of temporary soffit formwork and allowing rapid floor-by-floor construction. \todo[inline]{Add 1--2 sentences with a quantitative programme comparison if a source is available (e.g. typical floor cycle time for composite deck vs in-situ RC flat slab). SCI P300 or a contractor source may give this.}

\paragraph{Self-weight and foundation demand.} Although the composite slab is heavier than the timber systems, it is lighter than a solid RC flat slab of equivalent span for much of the range \todo[inline]{verify against final mass results --- Section~\ref{sec:res_office_depth} mass figure; state the span range over which composite is lighter than RC flat, if any, and qualify accordingly}, which reduces column and foundation loads and the seismic mass of the structure. In tall buildings this cumulative weight saving can be decisive and is not reflected in a per-floor GWP figure.

\paragraph{Structural depth and service integration.} \todo[inline]{Compare composite total construction depth (including services) against RC flat using the building-envelope results (Section~\ref{sec:res_envelope}). The argument: shallow composite floors, or the ability to pass services through deep-deck voids / cellular beams, can reduce storey height and therefore total building envelope, cladding area, and the embodied carbon of the rest of the structure --- a saving that offsets the higher slab GWP. State the depth comparison once the envelope analysis is final.}

\paragraph{Demountability and material recovery.} Steel decking and beams can in principle be unbolted and reused or recycled at end of life, whereas monolithic RC is typically downcycled to aggregate. The embodied-carbon accounting in this study is a cradle-to-gate (A1--A3) boundary \todo[inline]{confirm the LCA boundary actually used --- A1--A3 vs A1--A5 vs including module D; state explicitly} and therefore does not credit end-of-life recovery, which would partially close the gap to the RC systems.

In summary, the composite slab's higher cradle-to-gate GWP is a real and robust result, but it is one term in a multi-criteria decision. The contribution of this work is to quantify that term precisely and to show how it varies with span, occupancy, and continuity; the trade-offs above explain why a higher-GWP system continues to be selected, and motivate the cost and programme extensions proposed in Section~\ref{sec:disc_future}.

%==============================================================================
\subsection{Behaviour within the composite slab system}
\label{sec:disc_within}
%==============================================================================
\paragraph{Governing limit states.} The composite slab is stiffness-governed rather than strength-governed over most of the practical range: SLS deflection controls the single-span optima from \todo[inline]{insert governing span, ~5\,m}\,m upward, and the hogging checks control the continuous configurations (Section~\ref{sec:res_office_gov}). This has a direct GWP consequence---additional depth and material are being added to satisfy a serviceability limit, not to carry load---and explains why the optimum is insensitive to concrete grade (C25/30 is selected almost throughout, since a stronger grade adds GWP without adding the stiffness that governs).

\paragraph{Effect of span continuity.} The non-monotonic effect of continuity (Section~\ref{sec:res_office_continuity}) is one of the more nuanced findings. Continuity reduces composite GWP in the intermediate range, where the lower span moments allow a shallow deck to be retained, but penalises it at long spans, where the support hogging moment forces a deeper section and ultimately renders the two-span industrial case infeasible. This contrasts with the RC flat slab, which benefits from continuity across the whole range. \todo[inline]{Reference the continuity figure (fig:office\_continuity) and state the crossover span and the magnitude of the penalty at the longest feasible span, from final results.} The practical implication is that the benefit of designing a composite slab as continuous depends strongly on span, and that the simplified continuous-as-simply-supported approach of EN~1994-1-1 Cl.\,9.4.2(5) does not capture this, since it omits the hogging-governed regime that the explicit hogging checks in this model reveal.

\paragraph{Preferred deck profiles.} \todo[inline]{From the results, summarise which deck profiles are selected as optimal across the span range and why (shallow Multideck/ComFlor 60 at short spans, deep ComFlor 210/225 and Multideck 146 at long spans). Note the discrete jump in GWP at the shallow-to-deep deck transition and that this transition, not a smooth scaling, drives the GWP-span curve shape. This addresses the checklist item on comparing preferred deck profiles within the system.}

%==============================================================================
\subsection{Comparison between slab systems}
\label{sec:disc_between}
%==============================================================================
\paragraph{Ranking and application.} Across all scenarios the GWP ranking is consistent: timber rectangular lowest, then RC ribbed, RC flat, timber ribbed, and composite highest (Section~\ref{sec:res_office_gwp}). The ranking should, however, be read alongside the intended application of each system. \todo[inline]{Add a short paragraph or table listing typical applications: timber (low/medium-rise residential, low fire/acoustic demand, sustainability-led projects); RC flat (the default multi-storey solution, flexible layout, services below slab); RC ribbed (longer spans, material-efficient but formwork-intensive); composite (steel-framed multi-storey, fast construction, long spans, service integration). This is the checklist "list application uses for slab type" item and frames why a higher-GWP system is not simply discarded.}

\paragraph{Scope of the comparison.} An important caveat is that only the composite and RC flat systems were extended to continuous spans; the RC ribbed and timber systems are single-span analyses (Section~\ref{sec:res_intro}). The cross-system ranking is therefore most rigorous for the single-span case. For the continuous configurations the comparison is fair between composite and RC flat---the two systems most commonly used continuous in practice---but the ribbed and timber baselines should be read as single-span reference points rather than as fully optimised continuous designs. \todo[inline]{If time permits before submission, note whether extending the other systems to continuous spans is left as further work.}

\paragraph{Building-envelope perspective.} \todo[inline]{Once building\_envelope.py results are final (Section~\ref{sec:res_envelope}), discuss the total-height / building-envelope comparison within and across systems --- this is the checklist item "compare total height/building envelope within and across slab systems". Key point: a per-m2 slab comparison can mislead at the building scale if one system allows reduced storey height; reconcile the per-m2 ranking with the envelope-level ranking and state whether they agree.}

%==============================================================================
\subsection{Comparison with the literature}
\label{sec:disc_literature}
%==============================================================================
The central finding is consistent with the work of \cite{RefWorks:regulez2022sustainability}, who compared slab typologies including steel--concrete composites and concluded that the composite slab becomes progressively less efficient than reinforced concrete as span increases (Chapter~\ref{secondchapter}, Figure~\ref{fig:regulez2022}). The present results reproduce this trend and extend it: not only does the composite slab become relatively less efficient with span, it is the highest-carbon option across the entire practical range and all four occupancies, and the divergence is driven specifically by the SLS-deflection-governed deep-deck transition identified in Section~\ref{sec:res_office_depth}. \todo[inline]{If the Reguléz figures give numbers, add a quantitative comparison --- e.g. our composite-to-RC ratio at a given span vs theirs --- to show the results are of the same order, and discuss any divergence (different EPD sources, deck range, or design standard).}

A second point of contact with the literature concerns vertical shear. \cite{RefWorks:RefID:101-pereira2019contribution} show that the Eurocode~4 treatment of vertical shear neglects the contribution of the steel sheeting and is therefore conservative, since the real capacity of composite slabs is usually governed by longitudinal shear rather than vertical shear. In the present study vertical shear was never the governing check at any optimum, which is consistent with their finding; the conservatism in the code's vertical-shear treatment is therefore not a binding limitation here, but it does mean that the model, by following the code, slightly understates the true vertical-shear capacity. \todo[inline]{Confirm the Pereira \& Simões reference is the intended "vertical shear underestimate" study from the checklist; if a different RefWorks entry was meant, swap the citation. Optionally expand on whether crediting the sheeting would change any result (it would not, since vertical shear never governs).}

\todo[inline]{Optional: compare against \cite{RefWorks:broyles2024equations} (RC/PT systems) if a meaningful comparison can be drawn for the RC baselines.}

%==============================================================================
\subsection{Methodology: validity and limitations}
\label{sec:disc_method}
%==============================================================================
\paragraph{Verification against manufacturer data.} \todo[inline]{Populate once the manufacturer-table verification is final (Section~\ref{sec:res_validation}). Discuss: how closely the model's maximum permissible spans match the published load/span tables; where it is conservative or unconservative and why (e.g. the A\_pe = 0.9\,A\_p fallback, or the m--k vs partial-connection treatment); and what this implies for confidence in the GWP results. This is the checklist "verification results and discussion" item.}

\paragraph{EPD consistency and the ArcelorMittal exclusion.} The exclusion of ArcelorMittal decking on the grounds of a non-physical single-value EPD (Section~\ref{sec:deck_epd}) is a methodological strength rather than a gap: it prevents a systematic bias that would have favoured deep composite sections artificially. The reliance on Tata Steel and Kingspan EPDs, with KBOB as a reference, gives a consistent per-kilogram basis, but it also means the composite results are tied to two manufacturers' declared data; a different EPD dataset could shift the absolute composite GWP, though not enough to overturn the ranking given the size of the gap to the RC systems.

\paragraph{Vibration.} The slab-only vibration screening (Section~\ref{sec:vibration_method}) is conservative but does not capture full-floor behaviour; the worked example in Appendix~\ref{app:vibration} shows the full-floor frequency falling below the slab-alone value once beam flexibility is included. Vibration was not a binding constraint in the optimisation, but at long office spans the slab-alone frequency approaches the threshold, so a full-floor check would be warranted at detailed design.

\paragraph{End anchorage.} The omission of stud end anchorage (Section~\ref{sec:assumptions}, Appendix~\ref{app:end_anchorage}) is conservative and, as the appendix shows, affects only a handful of marginal sections at a non-negligible GWP cost; it does not alter the comparative conclusions.

\paragraph{Other simplifications.} \todo[inline]{Briefly consolidate the remaining limitations already flagged in the methodology: continuous-as-simply-supported design basis with explicit hogging checks; concrete-within-ribs neglected in self-weight (conservative); local resistance of sheeting at supports not checked; cradle-to-gate boundary. Cross-reference rather than repeat.}

%==============================================================================
\subsection{Concrete grade sensitivity (C20/25)}
\label{sec:disc_c20}
%==============================================================================
\todo[inline]{Populate from the dedicated C20/25 run (checklist "Research/Other work" item). The methodology restricts the optimisation to C25/30 and C30/37 on the grounds that C20/25 is not used in practice for these applications, yet C20/25 has the lowest GWP and would otherwise be selected. Report: for the office single-span case, the sections and governing checks chosen when C20/25 is permitted vs the C25/30 baseline, the GWP difference, and whether the governing check changes. Conclude whether the practice-based grade restriction materially affects the results. Run office scenario all-criteria with C20/25 alone, compare against the main case, then summarise here.}

%==============================================================================
\subsection{Swiss market context}
\label{sec:disc_swiss}
%==============================================================================
\todo[inline]{Populate from the "understand Swiss market availability of steel decking" research item. Points to cover: (1) the deck products in the database (Tata Steel ComFlor, Kingspan Multideck) are UK-market products; their availability in the Swiss market should be noted as a caveat on the direct applicability of the results to Swiss practice. (2) ArcelorMittal operates a Swiss-facing market and was excluded only on EPD-consistency grounds, not availability --- the same concept applies. (3) If decking is not readily available in Switzerland, a hollow-core ribbed alternative could serve the same structural role; note this as a Swiss alternative worth investigating. Keep factual and avoid over-claiming; flag any of these that need a source.}

%==============================================================================
\subsection{Further work}
\label{sec:disc_future}
%==============================================================================
Several extensions follow directly from the limitations above.

\paragraph{Cost analysis.} The most significant extension is a cost comparison alongside the GWP comparison. Because the systems differ greatly in construction method---composite construction trades higher material carbon for faster erection and reduced formwork---a cost and programme analysis would test whether the GWP penalty of composite construction is offset by economic and time savings, and would convert the multi-criteria argument of Section~\ref{sec:disc_why_composite} into quantified terms. \todo[inline]{State whether a costing framework or data source is identified; the checklist notes this as future work and suggests also indicating construction time per system.}

\paragraph{Low-carbon steel.} The composite GWP is dominated by the steel deck (Section~\ref{sec:res_office_breakdown}). Recycled-content or electric-arc-furnace (``green'') steel, with a markedly lower GWP intensity than the primary-route values used here, would reduce the composite GWP substantially and could narrow the gap to the RC systems. \todo[inline]{Quantify with a sensitivity estimate if a green-steel EPD value is available --- e.g. recompute the office composite GWP with a low-carbon deck intensity and state the resulting reduction. No green-steel reference is currently in the bibliography; one would need adding.}

\paragraph{End-of-life and module D.} Extending the LCA boundary to include end-of-life recovery (module D) would credit the recyclability and potential demountability of the steel components, and is expected to favour the composite and steel-containing systems relative to the cradle-to-gate comparison used here.

\paragraph{Continuous-span extension of all systems.} Extending the RC ribbed and timber systems to continuous spans would allow a fully consistent cross-system comparison in the continuous configurations, rather than the single-span baselines used at present.

\paragraph{Swiss-market decking and alternatives.} As noted in Section~\ref{sec:disc_swiss}, confirming Swiss-market availability of the deck products and investigating a hollow-core or alternative ribbed system for the Swiss context would improve the direct applicability of the results to Swiss practice.
